Local simple connectedness of boundaries of hyperbolic groups

### 2020-04-28 paper.tex
Removed todo. This will take more thought.
Henry's comments

\todo{Example of space ruled out by theorem}

\begin{theorem}\label{thm:main_theorem}
  Let $G$ be a one-ended hyperbolic group. Then $\boundary G$ is locally simply
  connected if and only if $\boundary G - \set\xi$ is simply connected for any
  point $\xi \in \boundary G$.
\end{theorem}

@@

@@

If $G$ is a group such that $\boundary G - \set\xi$ is disconnected for some
point $\xi$ in $\boundary G$, arguments due to Bowditch~\cite{bowditch98b} and
Swarup~\cite{swarup96} build an action of $G$ on a tree-like space to show that
$\boundary G$ is disconnected. It follows that $\boundary G$ is locally path
connected if and only if either $G$ is one- or two-ended. No such argument is
available in the higher dimensional case, but on the other hand no known
example of a boundary of a hyperblic group contradicts the corresponding
assertion. Therefore the following questions are natural.

\begin{question} Is there a one-ended hyperbolic group $G$ such that $\boundary
  G$ is simply connected but not locally simply connected? Conversely, if
  $\boundary G$ is locally simply connected but not simply connected, does it
  follow that $\boundary G$ is a circle, and so $G$ is virtually Fuchsian?
\end{question}

>>>>

\begin{theorem}\label{thm:main_theorem}
  Let $G$ be a one-ended hyperbolic group. Then $\boundary G$ is locally simply
  connected if and only if, for every point $\xi \in \boundary G$,  $\boundary
  G - \set\xi$ is simply connected.
\end{theorem}

@@

\subsection*{Acknowledgements}

I am grateful to Henry Wilton for pointing out the proof of
Proposition~\ref{prop:converse_theorem}, and for other helpful comments.  I am
also grateful to John Mackay and Panos Papazoglou for helpful discussions on
this topic.

@@

If $G$ is a group such that $\boundary G - \set\xi$ is disconnected for some
point $\xi$ in $\boundary G$, arguments due to Bowditch~\cite{bowditch98b} and
Swarup~\cite{swarup96} build an action of $G$ on a tree-like space to show that
$\boundary G$ is disconnected. It follows that $\boundary G$ is locally path
connected if and only if either $G$ is one- or two-ended. No such argument is
available in the higher dimensional case, but on the other hand no known
example of a boundary of a hyperblic group contradicts the corresponding
assertion. Therefore the following question is natural.

\begin{question} 
  Is there a one-ended hyperbolic group $G$ such that $\boundary G$ is simply
  connected but not locally simply connected? 
\end{question}

Note that the converse question has a positive answer if virtually Fuchsian
groups (which have boundary homeomorphic to a circle) are excepted:

\begin{proposition}\label{prop:converse_theorem}
  If $G$ is a one-ended hyperbolic group such that $\boundary G$ is locally
  simply connected then $\boundary G$ is either simply connected or
  homeomorphic to a circle.
\end{proposition}

\begin{proof}
  By Theorem~\ref{thm:main_theorem}, all point complements in $\boundary G$ are
  simply connected. Let $p$ and $q$ be distinct points in $\boundary G$. If
  $\boundary G - \set{p,q}$ is connected then $\boundary G$ is simply connected
  by the Seifert--van Kampen theorem. It follows that either $\boundary G$ is
  simply connected, or the complement of any two points in $\boundary G$ is
  disconnected, in which case $\boundary G$ is a circle
  by a classical theorem in point set topology~\cite[II.2.13]{wilder49}.
\end{proof}
